On the presentation of orthogonal groups over polynomial rings

### 2019-05-06 dl-paper.tex
Added Lemma 17

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(Xm)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. 
Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], M[X])$.

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r] & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\]
\end{lemma}
\begin{proof} 
We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type 3, 4 and 6.

@@

Finally, suppose $g$ has type $5$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon m$, $\eta=X$ and expressing $z_\alpha(f, mX)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, mX) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-mX^2 f) \cdot x_{-\beta}(mf) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon m) \cdot z_{-\beta}(-mf, -X) \cdot x_{-\alpha-\beta}(-\epsilon m^2X f) \cdot x_{-\alpha}(-m^2X^2 f), \end{multline}
and the latter expression is a product of generators of type 1, 1, 2, 5, 1, 2, 1, 4.
\end{proof}

For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From the above lemma it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

 As a side effect of this transformation, however, factors $x_\alpha(Xf)$ and $x_\beta(Xf)$ appearing in $g$ might be transformed into terms of the form
   $z_\alpha(Xf, X\xi)$, and $z_\beta(Xf, X\xi)$. Applying preceding lemma to the subsystem $\Psi$ we can write $\alpha = \alpha_1 - \alpha_2$ and 
    $\beta = \beta_1 - \beta_2$ with the additional property $\{ \pm \alpha, \pm \beta \} \cap \{ \alpha_1, \alpha_2, \beta_1, \beta_2 \} = \emptyset$.
 Invoking~\eqref{Z5} of~\cref{Zrels} with $\alpha = \alpha_1$, $\beta = -\alpha_2$, $\eta = 1$, $s = Xf$, $\xi = X\xi$ we can rewrite the term
  $z_\alpha(Xf, X\xi)$ as a product of generators of type 1,2,3. Similar argument also applies to $z_\beta(Xf, X\xi)$.

@@

>>>>

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(Xm)$, $m \in M$. The following lemma shows that this subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$. 
Invoking~\eqref{Z5} of~\cref{Zrels} with $s = Xf$, $\eta = 1$ we obtain that $z_{\pm\alpha}(Xf, \xi)$ also belong to $P_{\alpha}(*)$. Thus, we have shown that $P_\alpha(*)$ contains $\St(\Phi, A[X], XM[X])$.

@@

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r]^{p_\alpha} & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} 
We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type 3, 4 and 6.

@@

Finally, suppose $g$ has type $5$.
Substituting in~\eqref{Z5} of~\cref{Zrels} $s = -\epsilon f$, $\xi = -\epsilon m$, $\eta=X$ and expressing $z_\alpha(f, mX)$ through other terms we obtain that
\begin{multline} \nonumber z_\alpha(f, mX) = x_{\alpha+\beta}(\epsilon Xf) \cdot x_{\beta}(-mX^2 f) \cdot x_{-\beta}(mf) \cdot x_\alpha(f) \cdot \\ 
 \cdot z_{\alpha+\beta}(-\epsilon X f, -\epsilon m) \cdot z_{-\beta}(-mf, -X) \cdot x_{-\alpha-\beta}(-\epsilon m^2X f) \cdot x_{-\alpha}(-m^2X^2 f), \end{multline}
and the latter expression is a product of generators of type 1, 1, 2, 5, 1, 2, 1, 4.
\end{proof}
For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From the above lemma it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

 As a side effect of this transformation, however, factors $x_\alpha(Xf)$ and $x_\beta(Xf)$ appearing in $g$ might be transformed into terms of the form
   $z_\alpha(Xf, X\xi)$, and $z_\beta(Xf, X\xi)$. Applying preceding lemma to the subsystem $\Psi$ we can write $\alpha = \alpha_1 - \alpha_2$ and 
    $\beta = \beta_1 - \beta_2$ with the additional property $\{ \pm \alpha, \pm \beta \} \cap \{ \alpha_1, \alpha_2, \beta_1, \beta_2 \} = \emptyset$.
 Invoking~\eqref{Z5} of~\cref{Zrels} with $\alpha = \alpha_1$, $\beta = -\alpha_2$, $\eta = 1$, $s = Xf$, $\xi = X\xi$ we can rewrite the term
  $z_\alpha(Xf, X\xi)$ as a product of generators of type 1, 2, 3. Similar argument also applies to $z_\beta(Xf, X\xi)$.

@@

\begin{lemma}
 If $\alpha$ and $\beta$ form a sharp angle then one has $K(\alpha, \beta)^{x_{\beta - \alpha}(a)} \subseteq K(\alpha, \beta)$ for $a \in A$.
\end{lemma}
